---
title: "Homework assignment 1"
author: "Hussain, Mustaali"
format: pdf
---
```{r}
library(dplyr)
library(babynames)
```

# Question 2
```{r}
dim(babynames)
```
 There are 1924665 observations, and 5 variables

# Question 3
``` {r}
q3 <- tibble::tribble(
  ~variable_name, ~data_type, ~description,
  "year", "integer", "year baby was born",
  "sex", "categorical", "sex of the baby",
  "name", "character", "name of the baby",
  "n", "integer", "number of babies",
  "prop", "numeric", "proportion of the baby name within the year")
print(q3)

```

# Question 4
``` {r}
range(babynames$year)

```

The range is 1880 to  2017

# Question 5
``` {r}
babynames_no_n <- dplyr::select(babynames, year, sex, name, prop)

```

# Question 6


One reason for excluding n may be if we're only interested in comparing the relative popularity of names within a single year. Since the total number of births can vary greatly between years, prop is a better measure for comparing names within a year, because it normalizes for the total number of births. n on the other hand, is more suitable for comparing across years, where the total population might change. 




# Question 7a
``` {r}
second_millenium <- babynames_no_n %>% 
filter(year<= 2000) %>%
  group_by(sex) %>% 
  slice_max(prop, n=1) %>% 
  ungroup()

second_millenium

```
The most popular names within the 2nd millennium for males is John, and for females Mary

# Question 7b
``` {r}
babynames_no_n <- dplyr::select(babynames, year, sex, name, prop)
third_millenium <- babynames_no_n %>%
  filter(year >= 2001) %>% 
  group_by(sex) %>% 
  slice_max(prop, n=1) %>% 
  ungroup()

third_millenium

```
The most popular names within the 3rd millennium for males is Jacob, and for females Emily

# Question 8
```{r}
popular_names_qvx <- babynames %>%
  filter(year >= 2000, year <= 2012, grepl("^[QVX]", name)) %>%
  group_by(name) %>%
  summarize(total_n = sum(n)) %>%
  arrange(desc(total_n))
          
popular_names_qvx
          
```

The most popular names beginning with the letters Q, V, and X between 2000
and 2012 was Quinn, Victoria, and Xavier.

# Question 9
``` {r}
babynames_with_decade <- babynames %>%
  mutate(decade = floor(year / 10) * 10)
head(babynames_with_decade)



```

# Question 10
``` {r}
mean_median_babies <- babynames_with_decade %>%
  group_by(decade, sex) %>%
  summarize(
    mean_n = mean(n),
    median_n = median(n)
  ) %>%
  ungroup()
mean_median_babies
print(mean_median_babies, n = 28)

```

# Question 11a
``` {r}
mustaali_popularity <- babynames_with_decade %>%
  filter(name == "Mustaali") %>%
  slice_max(n, n = 1)


```


# Question 11b
```{r}
aimee_popularity <- babynames_with_decade %>%
  filter(name == "Aimee") %>%
  slice_max(n, n = 1)

```

# Question 11c
``` {r}
jack_popularity <- babynames_with_decade %>%
  filter(name == "Jack") %>%
  slice_max(n, n = 1)

scott_popularity <- babynames_with_decade %>%
  filter(name == "Scott") %>%
  slice_max(n, n = 1)

list(
  mustaali_popularity = mustaali_popularity,
  aimee_popularity = aimee_popularity,
  jack_popularity = jack_popularity,
  scott_popularity = scott_popularity
)
```

a) My name was not popular in any year or decade, as nobody from the data set has my name
b) My supervisor's name was most popular in the decade 1970, and the year 1977
c) The name Scott was most popular in the decade 1970, and the year 1971. The name Jack was most popular in the decade 1920, and the year 1927

