---
title: "Homework assignment 2"
author: "Hussain, Mustaali"
format: pdf
---
# Part A 
# Question 1 
``` {r}
library(tidyverse)
lordofrings_data<- readr::read_csv("lord-of-the-rings-trilogy.csv")
print(lordofrings_data)


```

# Question 2
- One rule of having tidy data is ensuring that every column is a single variable. However, each column from columns 2-7 contain multiple variables of both race and gender
- Also, there is inconsistency in the capitalization of words in the column names (ex: male and female are sometimes capitalized but not other times, hobbit is sometimes capitalized but not other times). This inconsistency interferes with having tidy data

# Question 3
- If this data set was tidy, it would have 4 columns and 18 rows. 
- When separating the columns into their own single variable, we would get 4 columns: "movie", "race", "gender", "word_count"
Since there are 3 races and 2 genders per race, each movie will have 6 combinations of race and gender (ex:elf female, elf male, hobbit female, hobbit male, man female, man male). And since their are  3 movies, this results in a total of 18 rows (6 rows per movie × 3 movies = 18 rows)

# Question 4 
- The column names in tidy format would be: "movie", "race", "gender", and "word_count"

# Part B

# Question 1
``` {r}
tidy_lordofrings_data <- lordofrings_data %>%
  pivot_longer(cols = elf_female:Man_male,
               names_to = c("race", "gender"),
               names_sep = "_",
               values_to = "word_count")

rename_with(tidy_lordofrings_data,tolower)
tidy_lordofrings_data <- tidy_lordofrings_data %>%
  mutate(race = tolower(race),
         gender = tolower(gender))

print(tidy_lordofrings_data)
```

# Question 2
``` {r}
library(dplyr)
male_hobbit_total <- tidy_lordofrings_data %>%
  filter(race == "hobbit", gender == "male") %>%
  summarize(total_word_count = sum(word_count))

female_elf_total <- tidy_lordofrings_data %>%
  filter(race == "elf", gender == "female") %>%
  summarize(total_word_count = sum(word_count))

male_elf_total <- tidy_lordofrings_data %>%
  filter(race == "elf", gender == "male") %>%
  summarize(total_word_count = sum(word_count))

print(male_hobbit_total)
print(female_elf_total)
print(male_elf_total)

```
a) the total number of words spoken by male hobbits is 8780 words
b) the total number of words spoken by female elves is 1743 words
c) the total number of words spoken by male elves is 1994 words

# Question 3
``` {r}
race_totals <- tidy_lordofrings_data %>%
  group_by(movie, race) %>%
  summarize(total_words = sum(word_count), .groups = 'drop')

total_words_per_movie <- race_totals %>%
  group_by(movie) %>%
  summarize(movie_total_words = sum(total_words), .groups = 'drop')

dominant_race <- race_totals %>%
  group_by(movie) %>%
  slice_max(total_words, n = 1) %>%
  left_join(total_words_per_movie, by = "movie") %>%
  mutate(dominance_proportion = total_words / movie_total_words)

print(dominant_race)

man_total_movie_1 <- tidy_lordofrings_data %>%
  filter(movie == "The Fellowship of the Ring", race == "man") %>%
  summarize(total_words = sum(word_count))

print(man_total_movie_1)


```

According to the generated tibble, in "The Fellowship of the Ring", the hobbit race accounts for 46.5% of all words spoken. In "The Return of the King", the man race accounts for 54.7% of all words spoken. In "The Two Towers", the man race accounts for 44.7% of all words spoken. Therefore, although the man race has the greater proportion of words spoken in 2 of the 3 movies, they do not necessarily "dominant", as their dominance proportion is still roughly at or under 50% of all words spoken in the 2 movies. In fact, if we calculate the proportion of words spoken by man across all 3 movies relative to all words spoken across all 3 movies, we get a proportion of ((1995+3990+2727)/(7853+7297+6095)) = 0.41 , or 41%, which indicates that man did not dominate in words spoken across all the 3 movies. 

# Question 4

``` {r}
race_totals <- tidy_lordofrings_data %>%
  group_by(movie, race) %>%
  summarize(total_words = sum(word_count), .groups = 'drop')

total_words_per_movie <- race_totals %>%
  group_by(movie) %>%
  summarize(movie_total_words = sum(total_words), .groups = 'drop')

dominant_race <- race_totals %>%
  group_by(movie) %>%
  slice_max(total_words, n = 1) %>%
  left_join(total_words_per_movie, by = "movie") %>%
  mutate(dominance_proportion = total_words / movie_total_words)

print(dominant_race)

```
-Yes, the dominant race does depend on the movie 
- In "The fellowship of the ring", we see that the dominant race in words spoken is hobbit, with 46.5% of total words spoken
- In both the  "The return of the king" and "The two towers" however, we see that the dominant race is man, with 54.7% and 44.7% of total words spoken, respectively
